An Empirical Analysis of C Preprocessor Use

### 2000-01-18 macros/papers/emp-use-2.tex
Insert the new paragraph, as tweaked by David and Mike.

Through extensive discussions, we have found that experts have widely
varying experiences with the preprocessor and similarly varying
expectations about what our data should and do show.  Some of our
specific analyses could have been performed differently, and some can
be (and have been) interpreted in various ways.
Our data, however, leave little doubt that the preprocessor is used in
broad and diverse ways, that most of those uses are simple, and that
there is certainly potential to reduce at least some of the complexity
introduced by the more sophisticated uses.

@@

  In \pkg{perl}, unknown identifiers make up 31\% of macro definitions,
  compared to an average of 6.9\% and \verb|ABOUT 9% NEED EXACT NUMBER| in
  \pkg{mosaic}, which has the the second largest proportion of unknown
  identifiers.  In order to avoid link-time collisions when the Perl
  interpreter is embedded in another application, \pkg{perl} conditionally
  redefines over one thousand global symbols to add the prefix
  \verb|Perl_|, as in {\tt \#define Xpv Perl_Xpv}.

@@

Finally, our results are of interest to software engineers researchers
for all of the above reasons and more.  Since this is the first study
of Cpp usage of which we are aware, it was worth performing simply to
determine whether the results were predictable \emph{a priori}; in
addition, we found some interesting, and in some cases surprising,
features of our suite of programs.

>>>>

Different programmers have different backgrounds and different tasks.  These
differences lead to substantial variations not only in how programmers use
the preprocessor, but also in their attitudes towards the preprocessor and
what data about preprocessor use they find of interest.  We
have found repeatedly that data that confirms one person's intuition comes
as a surprise to another\,---\,and every component of our data has had both
effects on different individuals.  Therefore, we provide a broad range
of analyses with the expectation that different readers will focus on
different parts, and will perhaps choose to extend our work in specific
directions.

@@

  In \pkg{perl}, unknown identifiers make up 31\% of macro definitions,
  compared to 9.0\% in \pkg{mosaic}, which has the the second largest
  proportion of unknown identifiers, and an average of 6.9\%.  In order to
  avoid link-time collisions when the Perl interpreter is embedded in
  another application, \pkg{perl} conditionally redefines over one thousand
  global symbols to add the prefix \verb|Perl_|, as in {\tt \#define Xpv
  Perl_Xpv}.

@@

Finally, our results are of interest to software engineering researchers
for all of the above reasons and more.  Since this is the first study
of Cpp usage of which we are aware, it was worth performing simply to
determine whether the results were predictable \emph{a priori}. 
Each aspect of our analysis has been surprising and interesting to
some individuals.  Pursuing more highly focused and deeper analyses
would be worthwhile.
